\singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants determine interactions with floral visitors

### 2022-06-13 Manuscript/Lanuza_et_al_2021.Rmd
edits

Flowering plants have an astonishing diversity of floral structures [@barrett2002; @schiestl2013] that shape plant-pollinator associations [@fenster2004; @dellinger2020]. However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints [@stearns1989]. Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits [@salguero2016; @friedman2020; @paterno2020; @roddy2021], their study is often limited to a handful number of reproductive traits that are rarely studied jointly. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other vegetative and physiological plant traits [@wright2004; @chave2009; @diaz2016; @onoda2017; @laughlin2021], there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the reproductive trait covariation patterns can help to further understand the different plant ecological strategies [@agrawal2020] and improve our knowledge on plant-pollinator associations [@roddy2021].

@@

Although the negative correlation between flower number and flower size has received substantial attention [worley2000; @sargent2007; @kettle2011], it has only been investigated across a few hundred of Angiosperm species with a two-trait approach. Here, we provide support to this evidence with a great diversity of phylogenetic lineages (170 families and over 1000 species) and show that other important traits like ovule number, style length and plant height also covariate along this reproductive axis.

quizas citar aqui el paper de maleness...

Hablar de los otros ejes ahora

The first axis described the negative association between flower number and flower size. Although this negative correlation has received substantial attention [worley2000; @sargent2007; @kettle2011], it has only been investigated across a few hundred of Angiosperm species and without a multitrait perspective. Here, we provide support to this evidence and show that also key reproductive traits to plant reproduction like ovule number and style length are positively correlated with flower size and negatively correlated with flower number. The second axis organized the different species by the degree of pollinator dependence where species with higher pollinator dependence showed greater floral display.

As expected... quizas hilar aqui con un par de refs para que quede menos results

@@

>>>>

Flowering plants have an astonishing diversity of floral structures [@barrett2002; @schiestl2013] that shape plant-pollinator associations [@fenster2004; @dellinger2020]. However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints [@stearns1989; @agrawal2020]. Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits [@salguero2016; @friedman2020; @paterno2020; @roddy2021], their study is often limited to a handful number of reproductive traits that are rarely studied jointly. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other vegetative and physiological plant traits [@wright2004; @chave2009; @diaz2016; @onoda2017; @laughlin2021], there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the reproductive trait covariation patterns can help to further understand the different plant ecological strategies [@agrawal2020] and improve our knowledge on plant-pollinator associations [@roddy2021].

@@

Although the negative correlation between flower number and flower size has received substantial attention [worley2000; @sargent2007; @kettle2011], it has only been investigated across a few hundred of Angiosperm species with a two-trait approach. Here, we provide support to this evidence with a great diversity of phylogenetic lineages (170 families and over 1000 species) and show that other reproductive and plant size and form related traits also covariate along this axis. For example, most species with unisexual flowers (14% of all species) were concentrated in the trait space and were associated with species with woddy life forms (i.e., trees and shrubs) with high flower number. This empirical evidence corroborates the idea that approximately 10% of species have separation of sexes [@barrett2002] and that breeding systems are associated with specific life forms [@petit2006]. The pollinator dependence - floral display axis

check number of species of goodwillie and what we offer that is new...

Further, the pollinator dependence - floral display axis provided additional support for the previously described positive association between higher outcrossing rate and larger floral display [@goodwillie2010].

Style length...

@@

[Plant scientists’ research attention is skewed towards colourful, conspicuous and broadly distributed flowers](May be interesting to comment on)

### 2022-06-13 Manuscript/Lanuza_et_al_2021.tex
edits

Flowering plants have an astonishing diversity of floral structures (\protect\hyperlink{ref-barrett2002}{Barrett, 2002}; \protect\hyperlink{ref-schiestl2013}{Schiestl \& Johnson, 2013}) that shape plant-pollinator associations (\protect\hyperlink{ref-dellinger2020}{Dellinger, 2020}; \protect\hyperlink{ref-fenster2004}{Fenster et al., 2004}). However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints (\protect\hyperlink{ref-stearns1989}{Stearns, 1989}). Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits (\protect\hyperlink{ref-friedman2020}{Friedman, 2020}; \protect\hyperlink{ref-paterno2020}{Paterno et al., 2020}; \protect\hyperlink{ref-roddy2021}{Roddy et al., 2021}; \protect\hyperlink{ref-salguero2016}{Salguero-Gómez et al., 2016}), their study is often limited to a handful number of reproductive traits that are rarely studied jointly. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other vegetative and physiological plant traits (\protect\hyperlink{ref-chave2009}{Chave et al., 2009}; \protect\hyperlink{ref-diaz2016}{Díaz et al., 2016}; \protect\hyperlink{ref-laughlin2021}{Laughlin et al., 2021}; \protect\hyperlink{ref-onoda2017}{Onoda et al., 2017}; \protect\hyperlink{ref-wright2004}{Wright et al., 2004}), there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the reproductive trait covariation patterns can help to further understand the different plant ecological strategies (\protect\hyperlink{ref-agrawal2020}{Agrawal, 2020}) and improve our knowledge on plant-pollinator associations (\protect\hyperlink{ref-roddy2021}{Roddy et al., 2021}).

@@

Although the negative correlation between flower number and flower size has received substantial attention {[}worley2000; \protect\hyperlink{ref-sargent2007}{Sargent et al.} (\protect\hyperlink{ref-sargent2007}{2007}); \protect\hyperlink{ref-kettle2011}{Kettle et al.} (\protect\hyperlink{ref-kettle2011}{2011}){]}, it has only been investigated across a few hundred of Angiosperm species with a two-trait approach. Here, we provide support to this evidence with a great diversity of phylogenetic lineages (170 families and over 1000 species) and show that other important traits like ovule number, style length and plant height also covariate along this axis.

quizas citar aqui el paper de maleness\ldots{}

Hablar de los otros ejes ahora

The first axis described the negative association between flower number and flower size. Although this negative correlation has received substantial attention {[}worley2000; \protect\hyperlink{ref-sargent2007}{Sargent et al.} (\protect\hyperlink{ref-sargent2007}{2007}); \protect\hyperlink{ref-kettle2011}{Kettle et al.} (\protect\hyperlink{ref-kettle2011}{2011}){]}, it has only been investigated across a few hundred of Angiosperm species and without a multitrait perspective. Here, we provide support to this evidence and show that also key reproductive traits to plant reproduction like ovule number and style length are positively correlated with flower size and negatively correlated with flower number. The second axis organized the different species by the degree of pollinator dependence where species with higher pollinator dependence showed greater floral display.

@@

@@

>>>>

Flowering plants have an astonishing diversity of floral structures (\protect\hyperlink{ref-barrett2002}{Barrett, 2002}; \protect\hyperlink{ref-schiestl2013}{Schiestl \& Johnson, 2013}) that shape plant-pollinator associations (\protect\hyperlink{ref-dellinger2020}{Dellinger, 2020}; \protect\hyperlink{ref-fenster2004}{Fenster et al., 2004}). However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints (\protect\hyperlink{ref-agrawal2020}{Agrawal, 2020}; \protect\hyperlink{ref-stearns1989}{Stearns, 1989}). Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits (\protect\hyperlink{ref-friedman2020}{Friedman, 2020}; \protect\hyperlink{ref-paterno2020}{Paterno et al., 2020}; \protect\hyperlink{ref-roddy2021}{Roddy et al., 2021}; \protect\hyperlink{ref-salguero2016}{Salguero-Gómez et al., 2016}), their study is often limited to a handful number of reproductive traits that are rarely studied jointly. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other vegetative and physiological plant traits (\protect\hyperlink{ref-chave2009}{Chave et al., 2009}; \protect\hyperlink{ref-diaz2016}{Díaz et al., 2016}; \protect\hyperlink{ref-laughlin2021}{Laughlin et al., 2021}; \protect\hyperlink{ref-onoda2017}{Onoda et al., 2017}; \protect\hyperlink{ref-wright2004}{Wright et al., 2004}), there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the reproductive trait covariation patterns can help to further understand the different plant ecological strategies (\protect\hyperlink{ref-agrawal2020}{Agrawal, 2020}) and improve our knowledge on plant-pollinator associations (\protect\hyperlink{ref-roddy2021}{Roddy et al., 2021}).

@@

Although the negative correlation between flower number and flower size has received substantial attention {[}worley2000; \protect\hyperlink{ref-sargent2007}{Sargent et al.} (\protect\hyperlink{ref-sargent2007}{2007}); \protect\hyperlink{ref-kettle2011}{Kettle et al.} (\protect\hyperlink{ref-kettle2011}{2011}){]}, it has only been investigated across a few hundred of Angiosperm species with a two-trait approach. Here, we provide support to this evidence with a great diversity of phylogenetic lineages (170 families and over 1000 species) and show that other reproductive and plant size and form related traits also covariate along this axis. For instance, most unisexual species (14\% of plant species of our dataset) were associated with woody species with high flower number. This concurs well with the previous idea that approximately 10\% of flowering plants have separation of sexes (\protect\hyperlink{ref-barrett2002}{Barrett, 2002}) and that species with unisexual flowers tend to be associated with woody/tree life forms (\protect\hyperlink{ref-petit2006}{Petit \& Hampe, 2006}).

The second axis organized the different species by the degree of pollinator dependence where species with higher pollinator dependence showed greater floral display.

@@

\href{May\%20be\%20interesting\%20to\%20comment\%20on}{Plant scientists' research attention is skewed towards colourful, conspicuous and broadly distributed flowers}

@@

\leavevmode\vadjust pre{\hypertarget{ref-petit2006}{}}%
Petit, R. J., \& Hampe, A. (2006). Some evolutionary consequences of being a tree. \emph{Annu. Rev. Ecol. Evol. Syst.}, \emph{37}, 187--214.
